### 2020-04-15 Thesis-Chapter2-Background.tex
Add stuff on US laws

If the weakness of data protection laws and absence of encryption protection laws left the door open for a second crypto
war, and increased government surveillance set events in motion toward conflict, it was the 2013 Snowden revelations
that constituted the crossing of the threshold. Though policies such as the U.S. \ac{NSA}'s warrantless wiretapping
caused a stir, it was the mass data collection under the agency's \ac{PRISM} and related programs that caused the real
public outcry \cite{landau_making_2013}. With the public interest focused on digital privacy, encryption promised a
technical solution. Under pressure from both the human rights field and from ordinary users demanding improved digital
privacy, U.S. tech companies responded by introducing default device encryption for \acl{DAR} and \acl{E2E} encryption
for \acl{DIM} \cite{treguer_us_2018}.

@@

The U.S. corporate response to Snowden should not be overly construed as a morally motivated defense of civil rights, as
their behavior is a matter of several practical factors \cite{treguer_us_2018}, not least of which is the pursuit of
basic digital security. Recall encryption and cryptography's critical role in security as described in
\mysec{sec-crypto-basics}, and that encryption's dual role as enabler of strong privacy and cornerstone of computer
security is a central conflict in the \ac{EA} debate. When questioned by lawmakers, technical leaders and executives
have repeatedly appealed to the fact that strong cryptography is a necessity for security \cite{schulze_clipper_2017}.

@@

% Amends Wiretap Act, Stored Communications Act, Pen Register Act. Deep history, but picking up here.

@@

@@

    \law{2001}{\acf{USAPATRIOT} Act}{sensenbrenner_2001}{

@@

    \law{2007}{\acf{PAA}}{mcconnell_s1927_2007}{

@@

    \law{2015}{\acf{USAFREEDOM} Act}{sensenbrenner_2015}{

@@

>>>>

If the weakness of data protection laws and absence of encryption protection laws left the door open for a second crypto
war, and increased government surveillance set events in motion toward conflict, it was the 2013 \ac{snowden}
revelations that constituted the crossing of the threshold. Though policies such as the U.S. \ac{NSA}'s warrantless
wiretapping caused a stir, it was the mass data collection under the agency's \ac{PRISM} and related programs that
caused the real public outcry \cite{landau_making_2013}. With the public interest focused on digital privacy, encryption
promised a technical solution. Under pressure from both the human rights field and from ordinary users demanding
improved digital privacy, U.S. tech companies responded by introducing default device encryption for \acl{DAR} and
\acl{E2E} encryption for \acl{DIM} \cite{treguer_us_2018}.

@@

The U.S. corporate response to \ac{snowden} should not be overly construed as a morally motivated defense of civil
rights, as their behavior is a matter of several practical factors \cite{treguer_us_2018}, not least of which is the
pursuit of basic digital security. Recall encryption and cryptography's critical role in security as described in
\mysec{sec-crypto-basics}, and that encryption's dual role as enabler of strong privacy and cornerstone of computer
security is a central conflict in the \ac{EA} debate. When questioned by lawmakers, technical leaders and executives
have repeatedly appealed to the fact that strong cryptography is a necessity for security \cite{schulze_clipper_2017}.

@@

Amended the Wiretap Act, Stored Communications Act, and Pen Register Act. Provided explicit protections for private
electronic communications data, and updated definitions to contemporary technology. It granted law enforcement explicit
access to certain data as well; some terminology has aged poorly with technological evolution, making the access
increasingly permissive \cite{shamsi_2011}.

@@

Mandated that telecommunications carriers enable the government to ``intercept all of the subscriber's wire and
electronic communications'' in a manner that ``protects the privacy and security'' of communications not authorized to
be intercepted. The law does not authorize law enforcement to compel use of any specific technologies, which meant that
the \ac{clipperchip} proposal issued a year earlier could not be made mandatory through \acs{CALEA}. \acs{CALEA}
specifically excludes ``information services'' from its interception requirement which has included internet platforms,
though that exclusion is currently under attack \cite{pfefferkorn_2019}.

@@

    \law{2001}{\ac{USAPATRIOTAct}}{sensenbrenner_2001}{

Written in the aftermath of September 11, the \ac{USAPATRIOTAct} broadly expanded U.S. government surveillance powers
in 25 sections of ``Title II: Enhanced Surveillance Procedures.'' The act notably expands \acs{FISA} order scope
(Section 206), duration (207), and use on U.S. citizens (218). Section 215 enables court orders requiring business
records, but the \ac{snowden} leaks revealed that the government had been secretly using a non-standard definition of
``business records'' in order ``to justify requests for domestic telephone metadata delivered in bulk, not
individualized requests'' \cite{landau_making_2013}.

@@

%     \law{2007}{\acf{PAA}}{mcconnell_s1927_2007}{
%
% Retroactively approved of GWB's warrentless wiretapping policy, a loose intepretation of FISA
% \cite{landau_making_2013}
%
% }

Like the \ac{USAPATRIOTAct} before it, \acs{FISAAA} expanded \ac{FISA} powers once again. It additionally provided
``telephone companies retroactive immunity for participating in the warrantless surveillance'' on any international
communication, a policy established in 2001 under another loose, secret interpretation of \ac{FISA}, ``as well as
creating prospective immunity for FISAAA activities'' \cite{landau_making_2013}.

@@

    \law{2015}{\ac{USAFREEDOMAct}}{sensenbrenner_2015}{

The \ac{USAFREEDOMAct}, introduced by the same U.S. House representative as the \ac{USAPATRIOTAct}, passed the senate
almost two years to the day after \ac{snowden}'s leaks began. It contains many reforms to \ac{FISA} and the use of
\acp{NSL}. Most notably, it ends the interpretation of section 215 of the \ac{USAPATRIOTAct} that enabled mass
collection of telephone metadata.

@@

\law{2019}{Technology in Criminal Justice Act}{demings_hr5227_2019}{

% pfefferkorn_2019
% And in fact, there’s already proposed legislation based on CSIS’s recommendations: the Technology in Criminal Justice
% Act of 2019, introduced in the House last month by a bipartisan group of co-sponsors led by Rep. Val Demings of Florida.
% The bill seeks to close the resource, training, and information-sharing gap for digital evidence, and to improve the
% frayed relationship between law enforcement and technology companies. Notably absent from the bill is any language that
% would roll back the CALEA compromise or mandate encryption backdoors. To be clear, I’m not advocating for or against
% Rep. Demings’ bill. I just want to point out that it offers an alternative to whatever backdoor nightmare the Senate
% Judiciary Committee is surely drafting up right now.

### 2020-04-15 Thesis-Glossary.tex
Add stuff on US laws

\newterm{disk-encryption}{disk encryption}{
    An application of ecnryption for \ac{DAR} that involves encrypting the contents
}
% End terms

@@

>>>>

\newterm{disk-encryption}{disk encryption}{
    An application of ecnryption for \ac{DAR} that involves encrypting the contents
}

\newterm{snowden}{Snowden}{
    Edward Snowden was an CIA contractor who leaked large volumes of data on NSA programs such as PRISM that performed
    mass collection of internet communications for surveillance
}
% End terms

@@

% Custom acronyms

\newabbreviation{USAPATRIOTAct}{USA PATRIOT Act}{Uniting and Strengthening America by Providing Appropriate Tools Required to Intercept and Obstruct Terrorism Act}
\newabbreviation{USAFREEDOMAct}{USA FREEDOM Act}{Uniting and Strengthening America by Fulfilling Rights and Ensuring Effective Discipline Over Monitoring Act}
